Data Migration under a Changing Schema

### 2023-07-30 2022Migration/articleMigrationFACS.tex
progress July 30th

%\def\id#1{\text{\it #1\/}}
\newcommand{\xrightarrowdbl}[2][]{%
  \xrightarrow[#1]{#2}\mathrel{\mkern-14mu}\rightarrow
}
\newcommand{\id}[1]{\text{\it #1\/}}
\newcommand{\code}[1]{\text{\tt\small #1}}
\newcommand{\stmtText}[1]{``{\small\tt #1}''}
\newcommand{\dom}[1]{\id{dom}(#1)}
\newcommand{\cod}[1]{\id{cod}(#1)}
%\renewcommand{\int}[2]{\id{inter}(#1,#2)}
\newcommand{\relsIn}[1]{\id{relsIn}(#1)}    % maps a Term to a set of Relations
\newcommand{\popF}[1]{\id{pop}_{#1}}
\newcommand{\pop}[2]{\popF{#1}(#2)}
\newcommand{\maintain}{\mathbin{\id{maint}}}
\newcommand{\enf}{\mathbin{{\tt enforce}}}
\newcommand{\enforce}[2]{{\tt enforce}_{#1}(#2)}
\newcommand{\instance}{\mathbin{\id{inst}}}
\newcommand{\relname}[1]{\id{relname}(#1)}
\newcommand{\evt}[2]{\id{event}_{#1,#2}}
\newcommand{\src}[1]{\id{src}(#1)}
\newcommand{\tgt}[1]{\id{tgt}(#1)}
\newcommand{\viol}[2]{{#1}(#2)}
%\newcommand{\violC}[1]{\id{viol}_{#1}}
\newcommand{\sign}[1]{\id{sign}(#1)}
\newcommand{\powerset}[1]{\cal{P}\{#1\}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\full}{V}
\newcommand{\declare}[3]{\id{#1}_{\pair{#2}{#3}}}
\newcommand{\subst}[3]{#3_{[#1\rightarrow #2]}}
\newcommand{\fullt}[2]{V_{\pair{#1}{#2}}}
\newcommand{\iden}{I}
\newcommand{\ident}[1]{I_{\id{#1}}}
\newcommand{\expr}[3]{(#1)_{#2\times #3}}
\newcommand{\pair}[2]{\langle{#1},{#2}\rangle}
\newcommand{\maprel}[2]{{\tt maprel}_{#1}({#2})}
\newcommand{\Pair}[2]{#1\times#2}
\newcommand{\pairs}[1]{\id{pairs}(#1)}
\newcommand{\triple}[3]{\langle{#1},{#2},{#3}\rangle}
\newcommand{\quadruple}[4]{\langle{#1},{#2},{#3},{#4}\rangle}
\newcommand{\atom}[1]{{\tt\small #1}}
\newcommand{\atoms}{\mathcal{A}}
\newcommand{\Atoms}{\mathbb{A}}
%\newcommand{\events}{\mathcal{E}}
%\newcommand{\Events}{\mathbb{E}}
\newcommand{\concept}[1]{{\tt\small #1}}
\newcommand{\concepts}{\mathcal{C}}
\newcommand{\Concepts}{\mathbb{C}}
\newcommand{\decls}{\mathcal{D}}  %% names of relations
\newcommand{\rels}{\mathcal{R}}   %% all relations
\newcommand{\Rels}{\mathbb{R}}   %% all relations
\newcommand{\relations}{\mathcal{M}} % representing terms. M is a subset of R.
\newcommand{\triples}{\mathcal{T}}
\newcommand{\Triples}{\mathbb{T}}
\newcommand{\Triple}[3]{#1\times#2\times#3}
\newcommand{\vertices}{N}
\newcommand{\enforces}{\mathcal{E}}
\newcommand{\rules}{\mathcal{U}}
\newcommand{\Rules}{\mathbb{U}}
\newcommand{\specrules}{\mathcal{S}}
\newcommand{\roles}{\mathcal{O}}
\newcommand{\dataset}{\mathscr{D}}
\newcommand{\Dataset}{\mathbb{D}}
\newcommand{\schema}{\mathscr{Z}}
\newcommand{\functionality}{\mathscr{F}}
\newcommand{\select}[2]{\id{select}_{#1}\{{#2}\}}
\newcommand{\migrsys}{\mathscr{M}}
\newcommand{\infsys}{\mathscr{S}}
\newcommand{\tf}[1]{\mathscr{T}(#1)}
\newcommand{\ptf}[1]{\mathscr{T}'(#1)}
\newcommand{\ti}[1]{\mathscr{I}(#1)}
\newcommand{\tic}[1]{I_{\cal C}(#1)}
\newcommand{\relAdd}{\dagger}
\newcommand{\flip}[1]{{#1}^\smallsmile} %formerly:  {#1}^\backsim
\newcommand{\kleeneplus}[1]{{#1}^+}
\newcommand{\kleenestar}[1]{{#1}^*}
\newcommand{\cmpl}[1]{\overline{#1}}
\newcommand{\rel}{\times}
\newcommand{\compose}{;}
\newcommand{\subs}{\subseteq}%{\models}
\newcommand{\fun}{\rightarrow}
\newcommand{\isa}{\preceq}
%\newcommand{\isaClos}{\sqsubseteq}
\newcommand{\typetest}{?}
\newcommand{\meet}{\sqcap}
\newcommand{\join}{\sqcup}
\newcommand{\Meet}{\bigsqcap}
\newcommand{\Moin}{\bigsqcup} % because LaTeX has already defined command \Join.
\newcommand{\order}{\ominus}
\newcommand{\anything}{\top}
\newcommand{\nothing}{\bot}
\newcommand{\rewriteto}{\rightarrow}
\newcommand{\calc}{\implies}
\newcommand{\alland}{\bigwedge}
\newcommand{\mph}[3]{#1_{#2\times #3}}
\newcommand{\mphu}[1]{#1_{\univ\times\univ}}

@@

   A relation $r$ can serve as a container of pairs,
   as defined by the function $\popF{r}:\Dataset\rightarrow\powerset{\Pair{\Atoms}{\Atoms}}$.
   It defines a set of pairs, also known as the population of $r$:
\begin{equation}
   \pop{r}{\dataset}\ =\ \{ \pair{a}{b}\mid\ \triple{a}{r}{b}\in\triples_{\dataset}\}
\label{eqn:pop-rel}
\end{equation}
%   Equation~\ref{eqn:wellTypedEdge} implies that for every data set $\dataset$:
%\[\pair{a}{b}\in\pop{\declare{n}{A}{B}}{\dataset}\ \Rightarrow\ a\instance_{\dataset}A\ \wedge\ b\instance_{\dataset}B\]
%   For a developer, this means that the type of an atom depends only on the relation in which it resides; not on the actual population of the database.
%
   We overload the notation $\popF{}$ so we can use it on concepts $\popF{C}:\Dataset\rightarrow\powerset{\Atoms}$
   and expressions.
\begin{align}
   \pop{C}{\dataset}&= \{ x\mid\ x\ \instance_{\dataset}\ C\}
\label{eqn:pop-concept}\\
   \pop{A - B}{\dataset}&= \pop{A}{\dataset} - \pop{B}{\dataset}
\label{eqn:pop-expr}
\end{align}

@@

   We describe a schema $\schema$ via a tuple $\quadruple{\concepts}{\rels}{\rules}{\enforces}$,
   in which $\concepts\subseteq \Concepts$ is a finite set of concepts,
   $\rels\subseteq \Rels$ is a finite set of relations,
   $\rules\subseteq {\Dataset \to \powerset{\Pair{\Atoms}{\Atoms}}}$ is a finite set of rules,
   and $\enforces\subseteq \Pair{\Rels}{\left(\Dataset \to \powerset{\Pair{\Atoms}{\Atoms}}\right)}$ is a finite set of enforcement rules.

   Each rule in a schema serves to constrain the data set at runtime, to ensure its semantic integrity.
   Every rule is an element of an infinite set called $\Rules$.
   In this paper, variables $u$ and $v$ represent rules.
   A rule $u$ in the schema is a function such that $\viol{u}{\dataset}$ represents any violations to $u$ in data set $\dataset$.
   If there are no violations, that is: $\viol{u}{\dataset} = \emptyset{}$, then we say that $u$ is satisfied.

@@

   Adding an enforcement rule to the script of a system specifies a function $u : \Dataset \to \powerset{\Pair{\Atoms}{\Atoms}}$
   with $\viol{u}{\dataset} = \pop{\id{term} - r}{\dataset}$.
   The system ensures that the rule is satisfied by adding a triple $\triple{x}{r}{y}$ for each $\pair{x}{y} \in \viol{u}{\dataset}$.
   Thus the system maintains $\pop{r}{\dataset} \supseteq \pop{\id{term}}{\dataset}$ as an invariant.

   Formally, enforcement rules are pairs with $u$ a function and $r$ a relation, written $r \mapsfrom u$ or equivalently $r \mapsfrom \lambda \dataset.~ u(\dataset)$.
   The set of enforcement rules in a schema is $\enforces$.

   \begin{definition}[Schema]
   A schema is a tuple $\quadruple{\concepts}{\rels}{\rules}{\enforces}$ that satisfies:
\begin{align}
   \declare{n}{A}{B}\in\rels&~\Rightarrow~ A\in\concepts\ \wedge\ B\in\concepts
   \label{eqn:relationsIntroduceConcepts}\\
   \pair{u}{r}\in\mathcal\enforces&~\Rightarrow~ r \in \rels\label{eqn:enforceRulesRels}
\end{align}
   \end{definition}
   Requirement~\ref{eqn:relationsIntroduceConcepts} ensures that concepts mentioned in a relation are defined in the schema.
   The type system of Ampersand~\cite{vdWoude2011} ensures that every system it generates satisfies this requirement.
   This is also known as static typing, which has well established advantages for the software engineering process~\cite{HanenbergKRTS14,Petersen2014}.

   %The notation $\schema_{\infsys}$ refers to the schema of system $\infsys$.
   When clarity is needed, we write $\concepts_{\schema}$, $\rels_{\schema}$, $\rules_{\schema}$, and $\enforces_{\schema}$
   for $\concepts$, $\rels$, $\rules$, and $\enforces$ corresponding to $\schema = \quadruple{\concepts}{\rels}{\rules}{\enforces}$.

\subsection{Information Systems}
\label{sct:Information Systems}
   Let us now define the notion of an information system.
\begin{definition}[information system]
\label{def:information system}
\item An information system $\infsys$ is a tuple $\pair{\dataset}{\schema}$, in which
\begin{itemize}
   \item $\dataset=\pair{\triples}{\instance}$ is a data set (satisfies equation~\ref{eqn:wellTypedEdge}), we write $\triples_\infsys = \triples$ and $\instance_\infsys = \instance$;
   \item $\schema=\quadruple{\concepts}{\rels}{\rules}{\enforces}$ is a schema (satisfies equation~\ref{eqn:relationsIntroduceConcepts} and~\ref{eqn:enforceRulesRels}), we write $\concepts_\infsys = \concepts$, $\rels_\infsys = \rels$, $\rules_\infsys=\rules$, and $\enforces_\infsys=\enforces$;
   \item All rules and enforcement rules are satisfied:
   \begin{align}
   \forall u\in\rules&.~\viol{u}{\dataset}=\emptyset\\
   \forall (r\mapsfrom u)\in\enforces&.~\viol{u}{\dataset} - \pop{r}{\dataset}=\emptyset
   \label{eqn:satisfaction}
   \end{align}
   \item Triples in the data set have their relation mentioned in the schema:
   \begin{eqnarray}
   \triple{a}{\declare{n}{A}{B}}{b}\in\triples&\Rightarrow&\declare{n}{A}{B}\in\rels
   \label{eqn:define R}
   \end{eqnarray}
\end{itemize}
\end{definition}
   % A \define{role} is a name that identifies a group of users.
   % It serves as a placeholder for a person or a machine (i.e. an actor) that works with the data set (i.e. create, read, update, or delete triples).
   % The purpose of a role is to mention an individual user (human) or an automated actor (bot) without knowing who that user is.

>>>>

%\def\id#1{\text{\it #1\/}}
\newcommand{\xrightarrowdbl}[2][]{%
  \xrightarrow[#1]{#2}\mathrel{\mkern-14mu}\rightarrow
}
\newcommand{\id}[1]{\text{\it #1\/}}
\newcommand{\code}[1]{\text{\tt\small #1}}
\newcommand{\stmtText}[1]{``{\small\tt #1}''}
\newcommand{\dom}[1]{\id{dom}(#1)}
\newcommand{\cod}[1]{\id{cod}(#1)}
%\renewcommand{\int}[2]{\id{inter}(#1,#2)}
\newcommand{\relsIn}[1]{\id{relsIn}(#1)}    % maps a Term to a set of Relations
\newcommand{\popF}[1]{\id{pop}_{#1}}
\newcommand{\pop}[2]{\popF{#1}(#2)}
\newcommand{\maintain}{\mathbin{\id{maint}}}
\newcommand{\enf}{\mathbin{{\tt enforce}}}
\newcommand{\enforce}[2]{{\tt enforce}_{#1}(#2)}
\newcommand{\instance}{\mathbin{\id{inst}}}
\newcommand{\relname}[1]{\id{relname}(#1)}
\newcommand{\evt}[2]{\id{event}_{#1,#2}}
\newcommand{\src}[1]{\id{src}(#1)}
\newcommand{\tgt}[1]{\id{tgt}(#1)}
\newcommand{\viol}[2]{\violC{#1}(#2)}
\newcommand{\violC}[1]{\id{viol}_{#1}}
\newcommand{\sign}[1]{\id{sign}(#1)}
\newcommand{\powerset}[1]{\cal{P}\{#1\}}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\full}{V}
\newcommand{\declare}[3]{\id{#1}_{\pair{#2}{#3}}}
\newcommand{\subst}[3]{#3_{[#1\rightarrow #2]}}
\newcommand{\fullt}[2]{V_{\pair{#1}{#2}}}
\newcommand{\iden}{I}
\newcommand{\ident}[1]{I_{\id{#1}}}
\newcommand{\expr}[3]{(#1)_{#2\times #3}}
\newcommand{\pair}[2]{\langle{#1},{#2}\rangle}
\newcommand{\maprel}[2]{{\tt maprel}_{#1}({#2})}
\newcommand{\Pair}[2]{#1\times#2}
\newcommand{\pairs}[1]{\id{pairs}(#1)}
\newcommand{\triple}[3]{\langle{#1},{#2},{#3}\rangle}
\newcommand{\quadruple}[4]{\langle{#1},{#2},{#3},{#4}\rangle}
\newcommand{\atom}[1]{{\tt\small #1}}
\newcommand{\atoms}{\mathcal{A}}
\newcommand{\Atoms}{\mathbb{A}}
%\newcommand{\events}{\mathcal{E}}
%\newcommand{\Events}{\mathbb{E}}
\newcommand{\concept}[1]{{\tt\small #1}}
\newcommand{\concepts}{\mathcal{C}}
\newcommand{\Concepts}{\mathbb{C}}
\newcommand{\decls}{\mathcal{D}}  %% names of relations
\newcommand{\rels}{\mathcal{R}}   %% all relations
\newcommand{\Rels}{\mathbb{R}}   %% all relations
\newcommand{\relations}{\mathcal{M}} % representing terms. M is a subset of R.
\newcommand{\triples}{\mathcal{T}}
\newcommand{\Triples}{\mathbb{T}}
\newcommand{\Triple}[3]{#1\times#2\times#3}
\newcommand{\vertices}{N}
\newcommand{\enforces}{\mathcal{E}}
\newcommand{\rules}{\mathcal{U}}
\newcommand{\Rules}{\mathbb{U}}
\newcommand{\specrules}{\mathcal{S}}
\newcommand{\roles}{\mathcal{O}}
\newcommand{\dataset}{\mathscr{D}}
\newcommand{\Dataset}{\mathbb{D}}
\newcommand{\schema}{\mathscr{Z}}
\newcommand{\functionality}{\mathscr{F}}
\newcommand{\select}[2]{\id{select}_{#1}\{{#2}\}}
\newcommand{\migrsys}{\mathscr{M}}
\newcommand{\infsys}{\mathscr{S}}
\newcommand{\tf}[1]{\mathscr{T}(#1)}
\newcommand{\ptf}[1]{\mathscr{T}'(#1)}
\newcommand{\ti}[1]{\mathscr{I}(#1)}
\newcommand{\tic}[1]{I_{\cal C}(#1)}
\newcommand{\relAdd}{\dagger}
\newcommand{\flip}[1]{{#1}^\smallsmile} %formerly:  {#1}^\backsim
\newcommand{\kleeneplus}[1]{{#1}^+}
\newcommand{\kleenestar}[1]{{#1}^*}
\newcommand{\cmpl}[1]{\overline{#1}}
\newcommand{\rel}{\times}
\newcommand{\compose}{;}
\newcommand{\subs}{\subseteq}%{\models}
\newcommand{\fun}{\rightarrow}
\newcommand{\isa}{\preceq}
%\newcommand{\isaClos}{\sqsubseteq}
\newcommand{\typetest}{?}
\newcommand{\meet}{\sqcap}
\newcommand{\join}{\sqcup}
\newcommand{\Meet}{\bigsqcap}
\newcommand{\Moin}{\bigsqcup} % because LaTeX has already defined command \Join.
\newcommand{\order}{\ominus}
\newcommand{\anything}{\top}
\newcommand{\nothing}{\bot}
\newcommand{\rewriteto}{\rightarrow}
\newcommand{\calc}{\implies}
\newcommand{\alland}{\bigwedge}
\newcommand{\mph}[3]{#1_{#2\times #3}}
\newcommand{\mphu}[1]{#1_{\univ\times\univ}}

@@

   A relation $r$ can serve as a container of pairs,
   as defined by the function $\popF{r}:\Dataset\rightarrow\powerset{\Pair{\Atoms}{\Atoms}}$.
   It defines a set of pairs, also known as the population of $r$:
\begin{equation}
   \pop{r}{\dataset}\ =\ \{ \pair{a}{b}\mid\ \triple{a}{r}{b}\in\triples_{\dataset}\}
\label{eqn:pop-rel}
\end{equation}
%   Equation~\ref{eqn:wellTypedEdge} implies that for every data set $\dataset$:
%\[\pair{a}{b}\in\pop{\declare{n}{A}{B}}{\dataset}\ \Rightarrow\ a\instance_{\dataset}A\ \wedge\ b\instance_{\dataset}B\]
%   For a developer, this means that the type of an atom depends only on the relation in which it resides; not on the actual population of the database.
%
   We overload the notation $\popF{}$ so we can use it on concepts $\popF{C}:\Dataset\rightarrow\powerset{\Atoms}$
   and expressions.
\begin{align}
   \pop{C}{\dataset}&= \{ x\mid\ x\ \instance_{\dataset}\ C\}
\label{eqn:pop-concept}\\
   \pop{x-y}{\dataset}&= \pop{x}{\dataset} - \pop{y}{\dataset}
\label{eqn:pop-expr}
\end{align}

@@

   We describe a schema $\schema$ via a tuple $\quadruple{\concepts}{\rels}{\rules}{\enforces}$,
   in which $\concepts\subseteq \Concepts$ is a finite set of concepts,
   $\rels\subseteq \Rels$ is a finite set of relations,
   $\rules\subseteq \Triple{(\Dataset \to \powerset{\Pair{\Atoms}{\Atoms}})}{\Concepts}{\Concepts}$ is a finite set of rules,
   and $\enforces\subseteq \Pair{\Rels}{\left(\Dataset \to \powerset{\Pair{\Atoms}{\Atoms}}\right)}$ is a finite set of enforcement rules.

   Each rule in a schema serves to constrain the data set at runtime, to ensure its semantic integrity.
   Every rule is an element of an infinite set called $\Rules$.
   In this paper, variables $u$ and $v$ represent rules.
   A rule $u$ in the schema is a triple $\triple{\violC{u}}{A}{B}$ such that $\viol{u}{\dataset}$ represents any violations to $u$ in data set $\dataset$.
   In Ampersand, rules are well-typed, just like relations:
   Every violation $\pair{a}{b}$ of a rule $\triple{\violC{u}}{A}{B}$ must satisfy $a\instance A$ and $b\instance B$.
   % For every data set $\dataset$, the set $\viol{u}{\dataset}$ is a subset of $\pop{A}{\dataset}\times\pop{B}{\dataset}$.
   If there are no violations, that is: $\viol{u}{\dataset} = \emptyset{}$, then we say that $u$ is satisfied.

@@

   Adding an enforcement rule to the script of a system specifies a function $\violC{u} : \Dataset \to \powerset{\Pair{\Atoms}{\Atoms}}$
   with $\viol{u}{\dataset} = \pop{\id{term} - r}{\dataset}$.
   The system ensures that the rule is satisfied by adding a triple $\triple{x}{r}{y}$ for each $\pair{x}{y} \in \viol{u}{\dataset}$.
   Thus the system maintains $\pop{r}{\dataset} \supseteq \pop{\id{term}}{\dataset}$ as an invariant.

   Formally, enforcement rules are pairs with $u$ a function and $r$ a relation, written $r \mapsfrom t$ or equivalently $r \mapsfrom \lambda \dataset.~ t(\dataset)$.
   The set of enforcement rules in a schema is $\enforces$.

   \begin{definition}[Schema]
   A schema is a tuple $\quadruple{\concepts}{\rels}{\rules}{\enforces}$ that satisfies:
\begin{align}
   \declare{n}{A}{B}\in\rels&~\Rightarrow~ A\in\concepts\,\wedge\, B\in\concepts
   \label{eqn:relationsIntroduceConcepts}\\
   \triple{\violC{u}}{A}{B}\in\rules&~\Rightarrow~A\in\concepts\,\wedge\, B\in\concepts\label{eqn:rule-has-type}\\
   \triple{\violC{u}}{A}{B}\in\rules&~\Rightarrow~\forall\dataset.\ \viol{u}{\dataset}\subseteq\Pair{\pop{A}{\dataset}}{\pop{B}{\dataset}}\label{eqn:rule-well-typed}\\
   (\declare{n}{A}{B}\mapsfrom t)\in\mathcal\enforces&~\Rightarrow~ \declare{n}{A}{B}\in \rels\label{eqn:enforcement-has-type}\\
   (\declare{n}{A}{B}\mapsfrom t)\in\mathcal\enforces&~\Rightarrow~\forall\dataset.\ \pop{t}{\dataset}\subseteq\Pair{\pop{A}{\dataset}}{\pop{B}{\dataset}}\label{eqn:enforcement-well-typed}
\end{align}
   \end{definition}
   Requirement~\ref{eqn:relationsIntroduceConcepts} ensures that concepts mentioned in a relation are defined in the schema.
   This is also known as static typing, which has well established advantages for the software engineering process~\cite{HanenbergKRTS14,Petersen2014}.
   Requirements~\ref{eqn:rule-has-type} and~\ref{eqn:rule-well-typed} ensure that violations in a rule correspond to the type of that rule. 
   Requirements~\ref{eqn:enforcement-has-type} and~\ref{eqn:enforcement-well-typed} ensure the same for enforcement rules. 
   The type system of Ampersand~\cite{vdWoude2011} ensures that every system it generates satisfies requirements~\ref{eqn:relationsIntroduceConcepts} thru~\ref{eqn:enforcement-well-typed}.
   When clarity is needed, we write $\concepts_{\schema}$, $\rels_{\schema}$, $\rules_{\schema}$, and $\enforces_{\schema}$
   for $\concepts$, $\rels$, $\rules$, and $\enforces$ corresponding to $\schema = \quadruple{\concepts}{\rels}{\rules}{\enforces}$.

\subsection{Information Systems}
\label{sct:Information Systems}
   Let us now define the notion of an information system.
\begin{definition}[information system]
\label{def:information system}
\item An information system $\infsys$ is a tuple $\pair{\dataset}{\schema}$, in which
\begin{itemize}
   \item $\dataset=\pair{\triples}{\instance}$ is a data set (satisfies equation~\ref{eqn:wellTypedEdge}), we write $\triples_\infsys = \triples$ and $\instance_\infsys = \instance$;
   \item $\schema=\quadruple{\concepts}{\rels}{\rules}{\enforces}$ is a schema (satisfies equation~\ref{eqn:relationsIntroduceConcepts} thru~\ref{eqn:enforcement-well-typed}), we write $\concepts_\infsys = \concepts$, $\rels_\infsys = \rels$, $\rules_\infsys=\rules$, and $\enforces_\infsys=\enforces$;
   \item All rules and enforcement rules are satisfied:
   \begin{align}
   \forall u\in\rules&.~\viol{u}{\dataset}=\emptyset\\
   \forall (r\mapsfrom t)\in\enforces&.~\viol{u}{\dataset} - \pop{r}{\dataset}=\emptyset
   \label{eqn:satisfaction}
   \end{align}
   \item Triples in the data set have their relation mentioned in the schema:
   \begin{eqnarray}
   \triple{a}{\declare{n}{A}{B}}{b}\in\triples&\Rightarrow&\declare{n}{A}{B}\in\rels
   \label{eqn:define R}
   \end{eqnarray}
\end{itemize}
\end{definition}
   % A \define{role} is a name that identifies a group of users.
   % It serves as a placeholder for a person or a machine (i.e. an actor) that works with the data set (i.e. create, read, update, or delete triples).
   % The purpose of a role is to mention an individual user (human) or an automated actor (bot) without knowing who that user is.
